Type-Safe Client-Server Communication in Web Applications

### 2014-02-26 paper.tex
Add stuff on unsafety in the face of unsafePerformIO/IO-dependent free vars.

>>>>

While our solution completely safe from the invocation point of the
\texttt{runApp} function onward, capturing IO-dependent variables introduced
before this invocation may lead to the client and server execution paths
diverging. Consider the following example:

\begin{verbatim}
main = do
  captured <- randomRIO (False, True)

  runApp config $ do
    diverged <- export $ \\transmitted -> do
      if captured == transmitted
        then return ``All OK''
        else return ``Diverged''

    runClient $ do
      divergence <- onServer (diverged <.> captured)
      alert divergence
\end{verbatim}

The dialog box displayed by this program will read ``All OK'' roughly half
of the times it is run, while in half of the runs it will read ``Diverged''.
The reason for this is that the \texttt{randomRIO} computation will be executed
twice - once on the client, and once on the server. Whenever the computation
returns differening results, the client and the server will have conflicting
opinions on the actual value of \texttt{captured}, causing any code
which - quite reasonably - assumes that variables are consistent between
client and server to fail.

This is easily remedied however, by a slight compiler modification. Instead of
using \texttt{main :: IO ()} as the program entry point, introduce a function
\texttt{appMain :: App Done} for this purpose, and let the runtime system start
the application by invoking \texttt{runApp config appMain}, where
\texttt{config} contains the URL and port configuration for the server, given
by the programmer using a compiler flag. As this modification only affects the
program's entry point, which is already given ``special'' treatment in Haskell
compilers, we believe this to be a practical, non-intrusive way to remove this
limitation, and in fact made this modification to the Haste compiler for our
implementation.

The astute reader may notice that the same applies to code using
\texttt{unsafePerformIO} to declare IO-dependent top-level values, and that
this problem is unaffected by changing the type of the program's entry point.
While this is indeed true, we feel that since it is quite possible to break
any Haskell program using \texttt{unsafePerformIO}, even subverting the type
system, this is not a problem with our solution.
